Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-29 introduction.tex
Edits to abstract, intro, and monitoring.

% Filtering
Recent studies have shown that filtering out monitoring checks on clean
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out both
propagation and checking of clean metadata. By keeping a separate set of flags
to track clean metadata, we are able to handle the common case of clean
metadata without the need to read full metadata. By propagating these flags, we
are also able to handle propagation of clean metadata efficiently in hardware.
Thus, we are able to have the flexibility of software-based monitors with
greatly reduced overheads.

% Adjustable overhead
With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. 
Traditionally, these overheads have been an all-or-nothing cost. If the
designer finds that the overheads exceed their desired budget, then monitoring
cannot be enabled.  
Thus, we present an architecture that enables a trade-off between overheads and
the monitoring coverage achieved. This is achieved by dynamically dropping
monitoring operations when overheads are too high. We prevent false
positives by extending the dataflow engine to also track these dropped
monitoring flows. We investigate three different policies on when dropping decisions are
made that cover a trade-off space between the ability to match a target overhead
and the monitoring coverage achieved.

@@

Thus, the main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
  \item We present an architecture that enables adjustable monitoring, creating a trade-off between monitoring coverage and overheads.
  \item We show how the dataflow tracking engine can be used to prevent false positives for adjustable monitoring.
  \item We explore the design space and trade-offs in dropping policies for performing adjustable monitoring.
\end{itemize}

>>>>

% Filtering
Recent studies have shown that filtering out monitoring checks on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out monitoring operations 
for null metadata. By keeping a separate set of flags
to track null metadata, we are able to handle the common case of null
metadata without the need to read the full metadata values. By propagating these flags, we
are also able to handle propagation of null metadata efficiently in hardware.
This filtering can be used to greatly reduce the overheads of monitoring.

% Adjustable overhead
There is a limit to how effective filtering is at reducing monitoring overheads. 
Traditionally, these overheads have been an all-or-nothing cost. If the
designer finds that the overheads exceed their desired budget, then monitoring
cannot be enabled.  
Thus, we propose to perform partial monitoring to enable further reduction of overheads.
 This is achieved by dynamically dropping
monitoring operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these dropped
monitoring flows. We investigate different policies for deciding which events
to drop for partial monitoring and show that there is a trade-off between
closely matching the overhead budget and the monitoring coverage achieved.

@@

The main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
  \item We present a hardware architecture using a dataflow tracking engine to efficiently enable filtering and partial monitoring.
  \item We investigate trade-off decisions for partial monitoring on the proposed architecture.
\end{itemize}

### 2014-05-29 monitoring.tex
Reorganized paper to unify discussion of dataflow engine.
Edits to abstract, intro, and monitoring.

\section{Parallel Run-Time Monitoring}
\label{sec:monitoring}

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:arch.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitoring core} performs a
series of \emph{monitoring operations}. The monitoring core operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Monitoring events are buffered in a FIFO structure to decouple the
running of the main core and the monitoring core. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitoring core may take several cycles to process a single event
from the main core. We refer to these stalls and other overheads, such as
contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base and bound addresses. On loads or stores with the
array
pointer, the monitoring core checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on certain ALU instructions to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

>>>>

\section{Parallel Programmable Run-Time Monitoring}
\label{sec:monitoring}

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:monitoring.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitoring core} performs a
series of \emph{monitoring operations}. The monitoring core operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
running of the main core and the monitoring core. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitoring core may take several cycles to process a single event
from the main core. We refer to these stalls and other overheads, such as
contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base (start) and bound (end) addresses. On loads or stores with the
array
pointer, the monitoring core checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on certain ALU instructions to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

### 2014-05-29 title.tex
Edits to abstract, intro, and monitoring.

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of computer
systems. However, these run-time monitors can introduce large performance and energy
overhead, especially when performed on programmable cores.
%Traditionally, these overheads have been an
%all-or-nothing cost; monitoring cannot be used if the overheads are considered
%too large.  
In this paper, we introduce a hardware dataflow tracking engine that can be
used to filter out unnecessary monitoring and enable partial monitoring.
The dataflow engine can identify events with clean metadata so that they can
be filtered out. To further reduce the overhead of monitoring, we propose
to enable a trade-off between monitoring coverage and overhead by dropping certain
monitoring operations. For the partial monitoring, the dataflow engine is used
to track dropped monitoring flows so that false positives can be avoided.
Given the architecture, we also investigate how the dropping decisions should be
made for partial monitoring and show that there exists a trade-off between closely
matching an overhead target and coverage that can be achieved.
% Our experiments show that the system is able to
% closely meet the target overhead and achieve significant monitoring coverage
% for reasonable overhead targets.
The experimental results show that filtering can significantly reduce monitoring
overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
XX\%-XX\%. Additionally, the partial monitoring is shown to still achieve XX\% coverage 
only with a 10\% overhead target.

>>>>

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of computer
systems. However, these run-time monitors can introduce large performance and energy
overheads, especially when performed on programmable cores.
%Traditionally, these overheads have been an
%all-or-nothing cost; monitoring cannot be used if the overheads are considered
%too large.  
In this paper, we introduce a hardware dataflow tracking engine that can be
used to filter out unnecessary monitoring and enable adjustable partial monitoring.
The dataflow engine can identify events with null metadata so that they can
be filtered out. To further reduce the overhead of monitoring, we propose
to enable a trade-off between monitoring coverage and overhead by dropping certain
monitoring operations. For this partial monitoring, the dataflow engine is used
to track dropped monitoring flows so that false positives can be avoided.
Given the architecture, we also investigate how the dropping decisions should be
made for partial monitoring and show that there exists a trade-off between closely
matching an overhead target and the coverage that can be achieved.
% Our experiments show that the system is able to
% closely meet the target overhead and achieve significant monitoring coverage
% for reasonable overhead targets.
The experimental results show that filtering can significantly reduce monitoring
overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
XX\%-XX\%. Additionally, the partial monitoring is shown to still achieve XX\% coverage 
only with a 10\% overhead target.
